---
title: "Semester-Project2"
author: "Shawn Vizgan"
date: "5/17/2021"
output: html_document
---

While it may not seem like the most important part of statistics, data visualization is actually super vital to the understanding of statistics by other scientists and everyday people alike. While the computation of various statistics is important work, the various forms of presentation can allow the products of sometimes complex designs and to be understood simply and quickly. While R has its own basic graphing tools, we will be looking at GGplot2 and it's various options.

Practical section


-How to use Bar Graphs
  -How to change colors
  -How to change the distance between bars
  
-How to use Line Graphs
  -how to change line colors
  =how to change the shape of points on the line
  
-Conceptual section
  -Why should someone use a bar graph versus a line graph (presentation of group means versus across time/conditions trends)
  -importance of showing variance scatter (random points on a graph or means in general can be deceiving and may not show the whole         story)
  -how to graph multiple independent variables
  
  
  In this example, the base functions for making graphs in R aren't that aesthetically pleasing.
  
```{r}
# load libraries
library(dplyr)
library(ggplot2)

# chickwts is a built in data set
# with chick weights by feed

# calculate means for each feed type
means_df <- chickwts %>%
  group_by(feed) %>%
  summarize(means = mean(weight))

# print table of means
knitr::kable(means_df)

plot(means_df)

```

Let's try putting this through ggplot2:

```{r}


# plot the means
ggplot(means_df, aes(x=feed, y=means, color = feed)) +
  geom_bar(stat="identity") +theme_minimal()










```


For linegraphs, “blank”, “solid”, “dashed”, “dotted”, “dotdash”, “longdash”, “twodash” are all options to choose from when it comes to how you want your line graph information to look. 


```{r}
df <- data.frame(dose=c("D0.5", "D1", "D2"),
                len=c(4.2, 10, 29.5))
head(df)

# Basic line plot with points
ggplot(data=df, aes(x=dose, y=len, group=1)) +
  geom_line()+
  geom_point()
# Change the line type
ggplot(data=df, aes(x=dose, y=len, group=1)) +
  geom_line(linetype = "dashed")+
  geom_point()
# Change the color
ggplot(data=df, aes(x=dose, y=len, group=1)) +
  geom_line(color="green")+
  geom_point()


```