---
title: "Semester-Project2"
author: "Shawn Vizgan"
date: "5/17/2021"
output: html_document
---

While it may not seem like the most important part of statistics, data visualization is actually super vital to the understanding of statistics by other scientists and everyday people alike. While the computation of various statistics is important work, the various forms of presentation can allow the products of sometimes complex designs and to be understood simply and quickly. While R has its own basic graphing tools, we will be looking at GGplot2 and it's various options.

Practical section


-How to use Bar Graphs
  -How to change colors
  -How to change the distance between bars
  
-How to use Line Graphs
  -how to change line colors
  =how to change the shape of points on the line
  
-Conceptual section
  -Why should someone use a bar graph versus a line graph (presentation of group means versus across time/conditions trends)
  -importance of showing variance scatter (random points on a graph or means in general can be deceiving and may not show the whole         story)
  -how to graph multiple independent variables
  
  
  In this example, the base functions for making graphs in R aren't that aesthetically pleasing.
  
```{r}

library(dplyr)
library(ggplot2)
means_df <- chickwts %>%
  group_by(feed) %>%
  summarize(means = mean(weight))


plot(means_df)

```

Let's try putting this through ggplot2, and specifically let's use a bar graph:

```{r}
library(ggplot2)



# plot the means
ggplot(means_df, aes(x=feed, y=means)) +
  geom_bar(stat="identity")

```
Bar graphs are one super common way to present your information. They are especially useful when highlighting differences between groups and categorical information. More info on bargraphs will be included in our conceptual section.These means come from one of the datasets available in ggplot2, but you can make your own data sets too for your bargraphs in R. There are many types of functions one can use to make data frames to base your graph around, and we will be using tibble. 


```{r}
library(tibble)
library(ggplot2)


# one factor three levels
fake_data <- tibble(IV1 = c("A","B","C"),
                    DV = c(3,6,5))


ggplot(fake_data, aes(x=IV1, y=DV))+
  geom_bar(stat='identity')

```
This is quite the barebones graph right now, but there are plenty of ways to spruce them up and modify what you're working with. One basic option is to add color. Within the aes part of the ggplot code, after adding in your x and y variables, you have the choice between "fill" and "color" which will affect your presentation's color choices. Fill corresponds to the inside of the bar whereas using only the color option changes just the bar's outline (it tripped me up the first time I encountered it too).

```{r}


ggplot(fake_data, aes(x=IV1, y=DV, fill=IV1))+
  geom_bar(stat='identity', color="black")

ggplot(fake_data, aes(x=IV1, y=DV, color=IV1))+
  geom_bar(stat='identity', color="green")


```
Thus, we have a basic working bargraph, but this may not be helpful for all designs and all forms of data. While we have our bars and their associated values, (and these are a good stand in for mean values we would see in the wild) not all data will be able to suit presentaion as bar graphs. Hence let us quickly present how to make line graphs in ggplot2 as well. That being said it is not as simple as just switching out geom_bar for geom_point as unlike in the bar graphs, line graphs need to be grouped together (hence the group=1 part of the code).

```{r}

ggplot(fake_data, aes(x=IV1, y=DV, group=1))+
  geom_point()+
  geom_line()

```

Some ways to liven up your bar graphs include changing the color and/or changing the way the lines look. Within the geom_point function you can play around with how the points look, whereas in geom_line you can change the colors and or line type as needed.
For your line graphs, “blank”, “solid”, “dashed”, “dotted”, “dotdash”, “longdash”, “twodash” are all options to choose from when it com es to how you want your line graph information to look. 

```{r}

ggplot(fake_data, aes(x=IV1, y=DV, group=1))+
  geom_point(shape = "triangle"
)+
  geom_line(color="blue")



ggplot(fake_data, aes(x=IV1, y=DV, group=1))+
  geom_point()+
  geom_line(linetype = "dashed")


```
Practical section:

The previous examples show some basic steps toward graph design, however a paltry 3 data points aren't enough towards representing how to model your data. Hence, here is how one should creathe multiple independent variables to graph, starting with just two. When making a balanced design, make sure to account for all separate conditions and counter balance them. For instance, we have 2 sets of 3 variables A, B, and C, and likewise we have 3 sets of alternating 1s and 2s.


``` {r}

fake_data <- tibble(IV1 = rep(c("A","B","C"), each=2),
                    IV2 = rep(c(1,2),3),
                    DV = c(3,6,5,3,4,6))

ggplot(fake_data, aes(x=IV1, y=DV, color = IV2))+
  geom_point()

```
When making barpgraphs, there is a choice between having stacked versus side by side versions of your bars. The choice between the two is yours, however typically to differentiate conditions better having them side by side is more common. Simply add position = dodged inside the geom_bar part of the function.

```{r}
ggplot(fake_data, aes(x=IV1, y=DV, fill=IV2))+
  geom_bar(stat='identity')

ggplot(fake_data, aes(x=IV1, y=DV, fill=IV2))+
  geom_bar(stat='identity', position="dodge")

```
When making a line graph, it is important to differentiate how you want your bar graph to identify its groupings between variable conditions, as otherwise it will simply look like random dots like in the points previously shown. How you code your colors whether inside a geom_line variable versus inside the aes function has different effects as well, depending on how you want to portray your data. It is especially advisable to add some sort of color or line style difference between graphs for the ease of view (unless you enjoy confusing your readers).

```{r}


# color depends on whether the IV is coded as a factor
fake_data <- tibble(IV1 = rep(c("A","B","C"), each=2),
                    IV2 = as.factor(rep(c(1,2),3)),
                    DV = c(3,6,5,3,4,6))

ggplot(fake_data, aes(x=IV1, y=DV, group=IV2, color = IV2))+
  geom_point()+
  geom_line()


```
It is also possible to code directly into the geom_line function instead of the aes function.This ends up grouping qualities around the lines themselves but not the points.



```{r}
ggplot(fake_data, aes(x=IV1, 
                      y=DV))+
  geom_point(shape = "square")+
  geom_line(aes(color=IV2,
                group=IV2,
                linetype=IV2))

```
Coding for 3 factors is similar to before, except we have to counter balance for yet another level of variables. It is also possible to hard code directly into tribble for how you want your design to look.Some final touches can be added such as adding your own title, adding a legend to differentiate variables.

```{r}

fake_data <- tribble(~IV1, ~IV2, ~IV3, ~DV,
                     "1A", "2A", "3A", 3,
                     "1A", "2A", "3B", 5,
                     "1A", "2B", "3A", 2,
                     "1A", "2B", "3B", 6,
                     "1B", "2A", "3A", 4,
                     "1B", "2A", "3B", 7,
                     "1B", "2B", "3A", 8,
                     "1B", "2B", "3B", 9,
                     "1C", "2A", "3A", 1,
                     "1C", "2A", "3B", 1,
                     "1C", "2B", "3A", 3,
                     "1C", "2B", "3B", 7
                     )

ggplot(fake_data, aes(x=IV1, 
                      y=DV, 
                      color=IV2,
                      group=IV2,
                      linetype=IV2,
                      shape=IV1))+
  geom_line(size=1)+
  geom_point(size=5, color="black")+
  geom_point(size=4)+
  theme_classic(base_size = 12)+
  ggtitle("the title of my graph") +
  xlab("x axis label") +
  ylab("y label") +
  coord_cartesian(ylim = c(-1,10))+
  geom_hline(yintercept = 0, color="grey")+
  facet_wrap(~IV3)

```


```
A final helpful conceptual feature to add would be on when to choose between bar graphs and line graphs. Line graphs are most commonly used when trying to portray data in ways to show trends across time, typically data that has some sort of fluctuation. Bar graphs on the other hand are used for group differences, mean comparisons, and showing off "facts." That being said, the same data can be portrayed as both line graphs and bar graphs depending on how you code the data, and for the most part it comes down to personal choice on which you choose. The specifc rules for bar graph are to use them for concrete rather than continuous variables (hence categories versus many times of day) for the sake of having a clear and comprehensive bar graph.






Practice Problems:

1) Make a line or bar graph for a dataset with 2 independent variables, with 2 different colors.

```{r}


New_data <- tibble(IV1 = rep(c("A","B","C"), each=2),
                    IV2 = as.factor(rep(c(1,2),3)),
                    DV = c(43,52,65,24,50,43))
ggplot(new_data, aes(x=IV1, y=DV, fill=IV2))+
  geom_bar(stat='identity', position="dodge")
  
  
  
  
  
  
```